Interval Probability for Fuzzy Quantum Theories

### 2016-06-18 prob.tex
prob.tex \section{Quantum Probability Spaces} updates...

In order to motivate the definition of quantum interval-valued probability
measures later, here is an equivalent definition of quantum probability
measures.

\begin{lemma}Given a Hilbert space $\Hilb$, a function~$\mu:\events\rightarrow[0,1]$
is a probability measure if and only if $\mu$ satisfies the following
conditions:
\begin{itemize}
\item $\mu(\mathbb{0})=0$, 
\item $\mu(\mathbb{1})=1$, 
\item $\mu\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P\right)=1\tensor*[_{\mathbb{R}}]{-}{}\mu\left(P\right)$,
and 
\item for mutually orthogonal projections $P_{i}$, we have $\mu\left(\tensor*[_{\mathscr{O}}]{\sum}{_{i}}P_{i}\right)=\tensor*[_{\mathbb{R}}]{\sum}{_{i}}\mu\left(P_{i}\right)$. 
\end{itemize}
\qed\end{lemma}

@@

For a given set of events $\events$, there are many possible probability
measures $\mu:\events\rightarrow[0,1]$. The Born rule~\cite{Born1983,Mermin2007,RiederSvozil2007},
a postulate of quantum mechanics, states that each pure quantum state
$\ket{\phi}$ induces a probability measure $\mu_{\phi}$ as follows:
\[
\mu_{\phi}(E)=\ip{\phi}{E\phi}
\]

Moreover, the Born rule can be extended to a mixed state. Suppose
we want to prepare a quantum system. As we discussed in classical
probability, our ability of preparing a state might not be prefect.
If we want to prepare $\ket{\phi}$, we may turn out preparing a set
of state~$\ket{\phi_{j}}$ each with probability~$q_{j}$. Then,
the state of the system can be expressed as a density matrix~$\rho=\sum_{j}q_{j}\proj{\phi_{j}}$,
where $\sum_{j}q_{j}=1$. It is natural that the quantum probability
measure introduced by $\rho$ is the combination of $\mu_{\phi_{j}}$
with respect to probability $q_{j}$~\cite{peres1995quantum,544199,RiederSvozil2007}:
\begin{eqnarray}
\mu_{\rho}\left(E\right) & = & \Tr\left(\rho E\right)=\sum_{j=1}^{N}q_{j}\mu_{\phi_{j}}\left(E\right)\textrm{ .}\label{BornRule.mixed}
\end{eqnarray}

Conversely, Gleason's theorem states that given a probability measure
$\mu$, there exist a mixed state~$\rho$ that induces such a measure
using the Born rule~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
The theorem is only valid in Hilbert spaces with dimension $d\geq3$.
It is instructive to study counterexamples in $d=2$, i.e., the case
of a one-qubit system.

\begin{example}[One-qubit quantum probability measure] Consider
a quantum probability measure~$\mu:\events\rightarrow[0,1]$ defined
as follow: 
\[
\mu(E)=\begin{cases}
1 & \textrm{, if }E=\proj{\ps}\textrm{ ;}\\
0 & \textrm{, if }E=\proj{\ms}\textrm{ ;}\\
\mu_{\ket{0}}(E) & \textrm{, otherwise.}
\end{cases}
\]
On one hand, $\mu$ is a probability measure. Because $\mu$ is almost
the same as a probability measure~$\mu_{\ket{0}}$, we only need
to check the orthogonal pair~$\proj{\ps}$ and $\proj{\ms}$: 
\[
\mu(\proj{\ps})+\mu(\proj{\ms})=1+0=1\textrm{ .}
\]
On the other hand, $\mu$ cannot be induced by any mixed state because
\[
\mu(\proj{\ps})=\mu(\proj{0})=1\textrm{ .}
\]
However, $\mu_{\rho}(E)=1$ if and only if $\rho$ represents a pure
state and $\rho=E$. \qed\end{example}

@@

>>>>

In order to motivate the definition of quantum interval-valued probability
measures later, we provide an equivalent definition of quantum probability
measures.

\begin{lemma}\label{lem:quantumProbabilityMeasure}Given a Hilbert
space $\Hilb$, a function~$\mu:\events\rightarrow[0,1]$ is a probability
measure if and only if $\mu$ satisfies the following conditions:
\begin{itemize}
\item $\mu(\mathbb{0})=0$, 
\item $\mu(\mathbb{1})=1$, 
\item $\mu\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P\right)=1\tensor*[_{\mathbb{R}}]{-}{}\mu\left(P\right)$,
and 
\item for mutually orthogonal projections $P_{i}$, we have $\mu\left(\tensor*[_{\mathscr{O}}]{\sum}{_{i}}P_{i}\right)=\tensor*[_{\mathbb{R}}]{\sum}{_{i}}\mu\left(P_{i}\right)$. 
\end{itemize}
\qed\end{lemma}

@@

For a given set of events $\events$, there are many possible probability
measures $\mu:\events\rightarrow[0,1]$. The Born rule~\cite{Born1983,Mermin2007,RiederSvozil2007},
a postulate of quantum mechanics, states that each pure quantum state
$\ket{\phi}$ induces a probability measure $\tilde{\mu}_{\phi}$
as follows: 
\[
\tilde{\mu}_{\phi}(E)=\ip{\phi}{E\phi}
\]

Moreover, the Born rule can be extended to a mixed state. Suppose
we want to prepare a quantum system. As we discussed in classical
probability, our ability of preparing a state might not be prefect.
If we want to prepare $\ket{\phi}$, we may turn out preparing a set
of state~$\ket{\phi_{j}}$ each with probability~$q_{j}$. Then,
the state of the system can be expressed as a density matrix~$\rho=\sum_{j}q_{j}\proj{\phi_{j}}$,
where $\sum_{j}q_{j}=1$. It is natural that the quantum probability
measure introduced by $\rho$ is the combination of $\tilde{\mu}_{\phi_{j}}$
with respect to probability $q_{j}$~\cite{peres1995quantum,544199,RiederSvozil2007}:
\begin{eqnarray}
\tilde{\mu}_{\rho}\left(E\right) & = & \Tr\left(\rho E\right)=\sum_{j=1}^{N}q_{j}\tilde{\mu}_{\phi_{j}}\left(E\right)\textrm{ .}\label{BornRule.mixed}
\end{eqnarray}

Conversely, in Hilbert spaces of dimension $d\geq3$, Gleason's theorem
states that given a probability measure $\mu$, there exist a mixed
state~$\rho$ that induces such a measure using the Born rule~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum}.
By applying Gleason's theorem, we can extend lemma~\ref{lem:quantumProbabilityMeasure}
to the following corollary.

\begin{cor}\label{cor:Gleason's}Given a Hilbert space $\Hilb$,
for any function~$\mu:\events\rightarrow[0,1]$ satisfies the conditions
listed in lemma~\ref{lem:quantumProbabilityMeasure}, there exist
an unique mixed state~$\rho$ such that $\mu=\tilde{\mu}_{\rho}$.\end{cor}

It is instructive to study counterexamples when $d=2$, i.e., the
case of a one-qubit system.

\begin{example}{[}One-qubit quantum probability measure{]} Consider
a quantum probability measure~$\mu:\events\rightarrow[0,1]$ defined
as follow: 
\[
\mu(E)=\begin{cases}
1 & \textrm{, if }E=\proj{\ps}\textrm{ ;}\\
0 & \textrm{, if }E=\proj{\ms}\textrm{ ;}\\
\mu_{\ket{0}}(E) & \textrm{, otherwise.}
\end{cases}
\]
On one hand, $\mu$ is a probability measure. Because $\mu$ is almost
the same as a probability measure~$\tilde{\mu}_{\ket{0}}$, we only
need to check the orthogonal pair~$\proj{\ps}$ and $\proj{\ms}$:
\[
\mu(\proj{\ps})+\mu(\proj{\ms})=1+0=1\textrm{ .}
\]
On the other hand, $\mu$ cannot be induced by any mixed state because
\[
\mu(\proj{\ps})=\mu(\proj{0})=1\textrm{ .}
\]
However, $\tilde{\mu}_{\rho}(E)=1$ if and only if $\rho$ represents
a pure state and $\rho=E$. \qed\end{example}

@@

Notice that if the last condition is replaced by equal sign, these
conditions will be exactly the same as the conditions in lemma~\ref{lem:quantumProbabilityMeasure}
which is equivalent to the Born rule as in corollary~\ref{cor:Gleason's}.
Therefore, it is reasonable to believe these conditions should be
equivalent to a interval-valued Born rule.
